### 2019-04-12 emsoft-2019/Error_Analysis.tex
section 4

%!TEX root = main.tex
\section{Error Analysis}
\label{sec:Error_Analysis}

We first present our error analysis assuming given fixed-point word lengths,
before explaining our optimization algorithm which determines suitable word
lengths fully automatically in~\autoref{sec:Controller_Synthesis}.
While we do not employ the error analysis independently of controller and
precision synthesis, we note that it is an alternative way of bounding errors to
existing methods~\cite{imperialrmpc}.

For simplicity of notation, we assume a uniform word length $p$ in this section,
though our algorithm (and implementation) support mixed-precision as well.

The output of an explicit MPC controller can be affected by two main sources
errors, which an error analysis has to capture\cite{imperialrmpc}:
\begin{itemize}
  \item[(i)] \emph{incorrect region selection}: 
    due to analog-digital conversion and quantization of $H_i$ and $K_i$,
    the controller can choose an incorrect control action $u_j$ instead of $u_i$

  \item[(ii)] \emph{approximate control output}: the control output is computed
    using quantized $F_i$ and $G_i$, accumulating errors.
\end{itemize}

We have to ensure that these errors remain below the error bound $\Delta$
used to design the robust MPC controller:
\begin{align}\label{eq:delta}
  \max_{\forall i,j. \mathit{neighbor}(\mathcal{R}_{i},\mathcal{R}_{j})}|u_{i}-u_{j}| + \max_{\forall i.\mathcal{R}_{i}\in\mathcal{R}} err(u_{i})_{p} \le \Delta
\end{align}
where $i,j$ are the indexes of two distinct regions $\mathcal{R}_{i}$ and
$\mathcal{R}_{j}$, \\ $\mathit{neighbor}(\mathcal{R}_{i},\mathcal{R}_{j})$
states that two regions share at least a point, $u_{i}$ is the control action
for controller $i$, and $err(u_{i})_{p}$ captures the approximation error for
computing the control action $u_{i}$ in fixed precision $p$.

\subsection{Incorrect Region Selection}

The first part of the error in~\autoref{eq:delta} captures the error due to
selecting of the wrong control action. That is, with infinite-precision arithmetic
and in the absence of any uncertainty from hardware measurements we would choose
control action $u_i$, but due to uncertainties we actually choose control action $u_j$:
\begin{align} \label{eq:maximization}
  \max_{\forall i,j.\mathit{neighbor}(\mathcal{R}_{i},\mathcal{R}_{j})}|u_{i}-u_{j}|&= \\
  \max_{\forall i,j.\mathit{neighbor}(\mathcal{R}_{i},\mathcal{R}_{j})}&|F_{i}\statevar+G_{i} - (F_{j}\statevar+G_{j})|\nonumber
\end{align}

%Because of the linearity of the function $u_{i}-u_{j}$, and because of the
%convexity of the regions $i$ and $j$, it is enough to evaluate function
%(\ref{eq:maximization}) at the corner points of the \texttt{tube}, instead of
%solving a maximization problem.

A naive computation of this error would consider all possible values of $x_k$.
This would lead to a gross over-approximation, however, since the controller
will choose a wrong region only in a relative narrow neighborhood, or
\emph{tube}, around the boundary between two regions $R_i$ and $R_j$.

The border between two regions $\mathcal{R}_{i}$ and $\mathcal{R}_{j}$ is defined by 
their corresponding boundaries:
\begin{equation}
\mathit{border}_{i,j} :=(H_i \statevar - K_i ) - (H_j \statevar - K_j)
\end{equation}
for appropriate matching borders $H_i, K_i$ and $H_j, K_j$.

In~\autoref{eq:maximization}, we thus constrain $\statevar$ to be within an appropriate tube:
\begin{equation}\label{eq:tube}
|\mathit{border}_{i,j}| \le \varepsilon
\end{equation}
where $\varepsilon$ captures the uncertainty in computing the
correct region and thus the width of the tube.
%
The width of the tube $\varepsilon$ is due to two sources of errors:
analog-to-digital conversion $\varepsilon_{A/D}$ and the quantization of region
bounds in memory $\varepsilon_{Q}$:
\begin{equation}\label{eq:epsilontot}
  \varepsilon=\varepsilon_{A/D}+\varepsilon_{Q}
\end{equation}

Analog conversion happens just before the controller receives the sensor input from the plant
and is given by:
\begin{equation*}
\varepsilon_{A/D}=\frac{V_{cc}}{2^{p}-1}
\end{equation*}
where $V_{cc}$ is the reference voltage of the converter (e.g. typically 5V), $p$
is the number of bits available to quantize the analog signal, and $2^{p}$ is
the resolution of the converter.

While $\varepsilon_{A/D}$ is intrinsic to the capabilities of the device,
$\varepsilon_{Q}$ depends on the precision used to store the boundaries:
\begin{equation}\label{eq:quantizationlines}
  \varepsilon_{Q} = |(H_i - \hat{H}_i) \qstatevar + (K_i - \hat{K}_i)|
\end{equation}
\eva{For which index is this evaluated, $i$, $j$ or both?}
That is, $\varepsilon_{Q}$ bounds the distance between any hyperplane in infinite
precision (H and K), and its counter-part quantized in $p$ bits precision
($\hat{H}_{p}$ and $\hat{K}_{p}$).
This second error can be tuned providing a trade-off between accuracy and
memory storage required.

\subsection{Approximate Control Output}

Once a control action $u_i$ is chosen, computing $u_{i}$ itself introduces
imprecision because $u_i$ needs to be evaluated in finite-precision arithmetic:
\begin{equation}\label{eq:fperror}
  err(u_{i})_{p} = | (F_i - \hat{F}_{i}) \statevar + ( G_i - \hat{G}_{i} ) |
\end{equation}
where $\hat{F}$ and $\hat{G}$ represent the quantized values (in $p$ bits) for
the infinite precision values $F$ and $G$.

\autoref{eq:fperror} is evaluated for all \statevarmath that are inside the
region $\mathcal{R}_i$, together with all the values in the tube surrounding
region $\mathcal{R}_i$.
In this way we compute the error also for those points that belong to a neighbor
of $\mathcal{R}_i$, but because of finite-precision errors they are erroneously
mapped to control action $u_i$.

\subsection{Implementation}

The incorrect region error (\autoref{eq:maximization}) is computed using
Matlab's multi-parametric toolbox. Due to the linearity of function
$u_{i}-u_{j}$, and because we defined the tube as a convex region, it suffices
to evaluate the function only at the corner points of the tube, and keep the
result with the maximal magnitude.
\eva{This needs more details. Which function(s) and algorithm from Matlab are used?}
\eva{How do you determine which regions are neighbors? resp. which borders are matching?}

$\varepsilon_Q$ and $err(u_{i})_{p}$ are computed by Daisy. For this we encode
the expressions in~\autoref{eq:quantizationlines} and \autoref{eq:fperror}
respectively as straight-line arithmetic expressions and specify the constraints
on the domain of $\statevar$ in the precondition.
Daisy performs a dataflow analysis to determine the finite-precision roundoff
errors. This error computation differs from the one employed by
\citet{imperialrmpc} which phrases the roundoff error computation as a (relaxed)
mixed integer programming problem.

While our actual synthesis algorithm (\autoref{sec:Controller_Synthesis}) does
not compute the errors $\varepsilon_Q$ and $err(u_{i})_{p}$ explicitly, the
error verification is used internally by Daisy to determine a suitable precision
$p$ and can also be called explicitly without the precision optimization.

% The approximation error and the precision tuning for both borders and active
% functions are computed in Daisy. $\varepsilon_{Q}$ is given in input to the
% analysis. After we compute the incorrect region in MATLAB we obtain the value
% for max error $err(u_{i})$ from (8). Then we give the errors and the vectors F,G
% and H,K in input to Daisy. We used Daisy as a verification tool: we provide in
% input the expression we want to quantize (e.g. $\mathcal{R}_{i}$) together with
% the error bound we can tolerate (e.g. $\varepsilon_{Q}$), and Daisy returns a
% precision configuration (uniform or mixed) such that the input error is
% satisfied. The way it finds the right configuration for uniform precision is the
% following: it starts by checking whether the input expression, quantized in
% unary uniform precision (1bit), introduces an error bounded by the input error.
% If this is not the case, Daisy increases the precision format by 1 bit and check
% whether the error now is satisfied. This unary increment runs inside a loop that
% terminates when the first valid format is matched. This uniform word-length is
% going to be the baseline for the mixed-precision tuning phase. The
% mixed-precision algorithm keeps a list of variables to quantize, and collects
% different candidate configurations that satisfy the input error. It returns the
% best configuration based on the evaluation of a quality function.