### 2019-06-11 emsoft-2019/V2/conclusion.tex
updates

\section{Conclusion}

We have described an automatic technique to tune an explicit MPC
controller that implements the controller in mixed-precision fixed-point arithmetic
while ensuring that the resulting loss of precision does not invalidate the
constraints of the original control problem.
Our technique models potential fixed-point imprecision explicitly as a disturbance
term, and uses robust explicit MPC to design a controller.
A static error analysis and a mixed-precision tuning tool is then 
used to find an efficient implementation of the controller function.
The implementation maintains the numerical error to within
the disturbance bounds used by the robust controller.
In experiments, we show that our technique can yield significant savings in memory,
ranging up to 20\%. 
In addition, the static analysis scales to controllers with hundreds of regions. \Mahmoud{All in all, using our proposed method, model predictive control under input and state constraints can be achieved while the memory footprint is minimized. By leveraging our method, memory requirement for implementation of explicit MPCs over the industrial embedded systems reduces. One exciting direction for future work would be to investigate effect of stochastic noises on the proposed design methodology.}